---
title: "Species specificity"
author: "Kent Riemondy RBI"
date: "`R Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: cosmo
    highlight: tango
    fig_caption: true
    code_folding: hide
    df_print: paged
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.width = 5)
knitr::opts_chunk$set(fig.height = 5)
```

```{r, message=F, warning=F, echo=F}
source("../../R/globals.R")
color_palette = c("#bdbdbd", "#1F78B4")
library("Seurat")
```

## mouse human mix

First examine our control scRNA-seq dataset, which is a 1:1 nih-3t3:293t mix. 

```{r}
mtx_dat <- read_tsv(file.path("..", "2018-05-16_mouse_human", 
                   "count_matrices", "original_10x_umis.tsv.gz"))

mtx_dat <- tidy_to_matrix(mtx_dat)
```

Calculate species specificity
```{r}

calc_species_purity <- function(mat, 
                               mouse_gene_prefix = "mm38::",
                               human_gene_prefix = "hg38::"){
  
  ## get mouse and human reads
  g_ids <- rownames(mat)
  mouse_ids <- str_subset(g_ids, str_c("^", mouse_gene_prefix))
  human_ids <- str_subset(g_ids, str_c("^", human_gene_prefix))
  
  mouse_counts <- colSums(mat[mouse_ids, ])
  human_counts <- colSums(mat[human_ids, ])
  counts <- data_frame(cells = colnames(mat),
                       mouse_counts,
                       human_counts)
  ## calculate species purity (human / human + mouse)
  res <- mutate(counts, 
                purity = human_counts / (human_counts + mouse_counts))
  res
}

cell_purity <- calc_species_purity(mtx_dat)

ggplot(cell_purity,
       aes(human_counts,
           mouse_counts)) +
  geom_point(aes(color = purity))
```

Next assign cells to species if > 80% UMIs are from 1 species

```{r assign_species}
cell_purity <- mutate(cell_purity,
                      species = ifelse(purity > 0.80, 
                                       "human",
                                       ifelse(purity < 0.20, 
                                              "mouse",
                                              "multiplet"))) 

ggplot(cell_purity,
       aes(human_counts,
           mouse_counts)) +
  geom_point(aes(color = species))
```

Now examine the specicies purity for the mouse and human cells

```{r species_purity}


ggplot(cell_purity, aes(species,purity)) +
  geom_boxplot(aes(fill = species))

```

## Downsample reads

Examine species specifity as number of reads are downsampled. Use umi flat file to downsample reads.

```{r}
umi_fn <- file.path(data_dir, 
                      "original_10x", 
                      "umis",
                      "umigroups.txt.gz")

umis <- read_tsv(umi_fn,
                 col_names = c("barcode_10x", 
                                 "umi_molecule", 
                                 "count")) %>% 
    filter(barcode_10x != "Cell_unmatched")

umis <- separate(umis, umi_molecule, c("umi_seq", "genome", "gene"), sep = "::")

ds_funs <- funs("downsampled_99" = rbinom(length(.), ., 0.01),
     "downsampled_90" = rbinom(length(.), ., 0.10),
     "downsampled_80" = rbinom(length(.), ., 0.20),
     "downsampled_70" = rbinom(length(.), ., 0.30),
     "downsampled_60" = rbinom(length(.), ., 0.40),
     "downsampled_50" = rbinom(length(.), ., 0.50),
     "downsampled_40" = rbinom(length(.), ., 0.60),
     "downsampled_30" = rbinom(length(.), ., 0.70),
     "downsampled_20" = rbinom(length(.), ., 0.80),
     "downsampled_10" = rbinom(length(.), ., 0.90),
     "downsampled_0" = rbinom(length(.), ., 1.00))
                    
umis_ds <- mutate_at(umis, .vars = "count", .funs = ds_funs)
```

```{r plt_summaries}
cols_to_summarize <- str_subset(colnames(umis_ds), "downsampled")
names(cols_to_summarize) <- cols_to_summarize

ds_umi_summaries <- map_dfr(cols_to_summarize,  
                            function(x) {
                              x <- sym(x)
                              filter(umis_ds, !!x > 0) %>% 
                              group_by(barcode_10x) %>% 
                              summarize(human = sum(genome == "hg38"),
                                mouse = sum(genome == "mm38"),
                                purity = human / (human + mouse))},
                            .id = "downsampling")

ds_read_summaries <- map_dfr(cols_to_summarize,  
                            function(x) {
                              x <- sym(x)
                              umis_ds %>% 
                                filter(!!x > 0) %>% 
                                group_by(barcode_10x, genome) %>% 
                                summarize(counts = sum(!!x)) %>% 
                                spread(genome, counts) %>% 
                                mutate(purity = hg38 / (hg38 + mm38)) %>% 
                                select(barcode_10x, 
                                       mouse = mm38,
                                       human = hg38,
                                       purity)},
                            .id = "downsampling")

ds_summaries <- bind_rows(list(
  UMIs = ds_umi_summaries,
  Reads = ds_read_summaries), .id = "type")

ds_summary <- inner_join(ds_summaries, 
           cell_purity[, c("cells", "species")], 
           by = c("barcode_10x" = "cells")) %>% 
  separate(downsampling, c("id", "amount"), sep = "_") %>% 
  mutate(species = str_to_title(species))

x_axis_labels <- sort(unique(ds_summary$amount))

reordered_labels <- x_axis_labels[order(as.numeric(x_axis_labels), decreasing = T)]

ds_summary <- ds_summary %>% 
  mutate(downsampling_percentage = factor(amount,
                                          levels = reordered_labels))

ds_summary <- filter(ds_summary, species != "Multiplet")

ds_summary <- mutate(ds_summary, 
                     adjusted_purity = ifelse(species == "Mouse",
                                              1 - purity,
                                              purity))
p <- ggplot(ds_summary, 
       aes(downsampling_percentage, adjusted_purity)) +
  geom_boxplot(aes(fill = type),
               coef = Inf) +
  facet_wrap(~species, scales = "free_y") +
  scale_fill_brewer(palette = "Set1",
                    name = "") +
  labs(x = "% downsampling",
       y = "Species purity") +
  theme(legend.position = "top") 

p

save_plot("species_spec_umi_vs_reads.pdf", p,
          base_aspect_ratio = 1.5)
```


## Downsample reads from 10x dataset

Examine species specifity as number of reads are downsampled. Use umi flat file to downsample reads.

```{r 10x}

barcodes <- dir(file.path(data_dir, "public_10x_mh", "fastq", "original"),
                pattern = "*_*.tsv", full.names = T)
names(barcodes) <- basename(barcodes) %>% str_split(., "_", simplify = T) %>% .[, 1]
barcodes <- map_dfr(barcodes, read_tsv, col_names = "barcode_10x", .id = "species")

umi_fn <- file.path(data_dir, 
                      "public_10x_mh", 
                      "umis",
                      "umigroups.txt.gz")

umis <- read_tsv(umi_fn,
                 col_names = c("barcode_10x", 
                                 "umi_molecule", 
                                 "count")) %>% 
    filter(barcode_10x != "Cell_unmatched")

umis <- separate(umis, umi_molecule, c("umi_seq", "genome", "gene"), sep = "::")

ds_funs <- funs("downsampled_99" = rbinom(length(.), ., 0.01),
     "downsampled_90" = rbinom(length(.), ., 0.10),
     "downsampled_80" = rbinom(length(.), ., 0.20),
     "downsampled_70" = rbinom(length(.), ., 0.30),
     "downsampled_60" = rbinom(length(.), ., 0.40),
     "downsampled_50" = rbinom(length(.), ., 0.50),
     "downsampled_40" = rbinom(length(.), ., 0.60),
     "downsampled_30" = rbinom(length(.), ., 0.70),
     "downsampled_20" = rbinom(length(.), ., 0.80),
     "downsampled_10" = rbinom(length(.), ., 0.90),
     "downsampled_0" = rbinom(length(.), ., 1.00))
                    
umis_ds <- mutate_at(umis, .vars = "count", .funs = ds_funs)
```

```{r plt_summaries_10x, fig.cap= "% of reads aligned to each species for mouse or human cells. Fastq data was downloaded from the 10x genomics website and processed through the same pipeline as our data."}
cols_to_summarize <- str_subset(colnames(umis_ds), "downsampled")
names(cols_to_summarize) <- cols_to_summarize

ds_umi_summaries <- map_dfr(cols_to_summarize,  
                            function(x) {
                              x <- sym(x)
                              filter(umis_ds, !!x > 0) %>% 
                              group_by(barcode_10x) %>% 
                              summarize(human = sum(genome == "hg38"),
                                mouse = sum(genome == "mm38"),
                                purity = human / (human + mouse))},
                            .id = "downsampling")

ds_read_summaries <- map_dfr(cols_to_summarize,  
                            function(x) {
                              x <- sym(x)
                              umis_ds %>% 
                                filter(!!x > 0) %>% 
                                group_by(barcode_10x, genome) %>% 
                                summarize(counts = sum(!!x)) %>% 
                                spread(genome, counts) %>% 
                                mutate(purity = hg38 / (hg38 + mm38)) %>% 
                                select(barcode_10x, 
                                       mouse = mm38,
                                       human = hg38,
                                       purity)},
                            .id = "downsampling")

ds_summaries <- bind_rows(list(
  UMIs = ds_umi_summaries,
  Reads = ds_read_summaries), .id = "type")

ds_summary <- inner_join(ds_summaries, 
           barcodes, 
           by = "barcode_10x") %>% 
  separate(downsampling, c("id", "amount"), sep = "_") %>% 
  mutate(species = str_to_title(species))

x_axis_labels <- sort(unique(ds_summary$amount))

reordered_labels <- x_axis_labels[order(as.numeric(x_axis_labels), decreasing = T)]

ds_summary <- ds_summary %>% 
  mutate(downsampling_percentage = factor(amount,
                                          levels = reordered_labels))



ds_summary <- mutate(ds_summary, 
                     adjusted_purity = ifelse(species == "Mouse",
                                              1 - purity,
                                              purity))
ds_summary <- filter(ds_summary, adjusted_purity > 0.80)

p <- ggplot(ds_summary, 
       aes(downsampling_percentage, adjusted_purity)) +
  geom_boxplot(aes(fill = type),
               coef = Inf) +
  facet_wrap(~species, scales = "free_y") +
  scale_fill_brewer(palette = "Set1",
                    name = "") +
  labs(x = "% downsampling",
       y = "Species purity") +
  theme(legend.position = "top") 
p

save_plot("species_spec_umi_vs_reads_10xdata.pdf", p,
          base_aspect_ratio = 1.5)
```
