---
title: "process 10k pbmcs"
author: "Kent Riemondy RBI"
date: "6/1/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```


```{r, message=F, warning=F, echo=F}
source("../../R/globals.R")
```

```{r, eval = F}
file_url <- "http://cf.10xgenomics.com/samples/cell-exp/1.1.0/fresh_68k_pbmc_donor_a/fresh_68k_pbmc_donor_a_filtered_gene_bc_matrices.tar.gz"
dir.create("10x_data")
download.file(
  file_url, file.path("10x_data", "pbmc68k.tar.gz")
)
untar(file.path("10x_data", "pbmc68k.tar.gz"))
unlink(file.path("10x_data", "pbmc68k.tar.gz"))
```

```{r}
library(Seurat)
```

```{r, eval = F}
sobj <- CreateSeuratObject(Read10X(file.path("filtered_matrices_mex", 
                                             "hg19")),
                           min.genes = 200)

sobj <- NormalizeData(sobj)
sobj <- FindVariableGenes(sobj, do.plot = T, y.cutoff = 0.75)
```


Only run scaling, PCA, tSNA, and clustering using variable genes in dataset. Otherwise the scaling will consume too much memory (40Gb on my laptop before I killed it).

```{r figure_out_pcs, eval = F}

sobj <- ScaleData(sobj, genes.use = sobj@var.genes)

sobj <- RunPCA(sobj, 
               pc.genes = sobj@var.genes, 
               pcs.compute = 20, 
               do.print = F, seed.use = 20180525)
PCElbowPlot(sobj)
```

```{r, eval = F}
sobj <- RunTSNE(sobj, dims.use = 1:20, perplexity = 50, seed.use = 20180525)
TSNEPlot(sobj)
FeaturePlot(sobj, c("PPBP", "PF4"))
```


```{r, eval = F}
sobj <- FindClusters(sobj,
                     dims.use = 1:20, 
                     k.param = 30,
                     n.start = 10,
                     nn.eps = 0.5,
                     resolution = 1.2, 
                     print.output = F, 
                     random.seed = 20180525)
```

```{r, eval = F}
TSNEPlot(sobj, do.label = T, colors.use = c(brewer.pal(12, "Paired"),
                                            brewer.pal(9, "Set1")))

sobj@meta.data %>% 
  group_by(res.1.2) %>% 
  summarize(n = n())
```

Find markers for MK cluster (marked by PF4, PPBP). Need to downsample the non megakaryocyte comparision to avoid an error (Error in asMethod(object) : 
  Cholmod error 'problem too large' at file ../Core/cholmod_dense.c, line 105), due to an error with subsetting a sparse matrix to normal one in the FindMarkers code (https://github.com/satijalab/seurat/issues/417). 

Actually downsampling does nothing, because the error occurs prior to the downsampling. Here is a hacked version of the seurat function that works. Consider submitting a PR. 

```{r, eval = F}
saveRDS(sobj, "sobj.rds")
```

```{r}
sobj <- readRDS("sobj.rds")
```

```{r}
CustomFindMarkers <- function(
  object,
  ident.1,
  ident.2 = NULL,
  genes.use = NULL,
  logfc.threshold = 0.25,
  test.use = "wilcox",
  min.pct = 0.1,
  min.diff.pct = -Inf,
  print.bar = TRUE,
  only.pos = FALSE,
  max.cells.per.ident = Inf,
  random.seed = 1,
  latent.vars = NULL,
  min.cells.gene = 3,
  min.cells.group = 3,
  pseudocount.use = 1,
  assay.type = "RNA",
  ...
) {
  data.use <- object@data
  genes.use <- rownames(x = data.use)

  # in case the user passed in cells instead of identity classes
  cells.1 <- WhichCells(object = object, ident = ident.1)
  
  # if NULL for ident.2, use all other cells
  if (length(x = as.vector(x = ident.2) > 1) && any(as.character(x = ident.2) %in% object@cell.names)) {
    cells.2 <- intersect(x = ident.2, y = object@cell.names)
  } else {
    if (is.null(x = ident.2)) {
      # cells.2 <- object@cell.names
      cells.2 <- Seurat::WhichCells(object = object,
                                    cells.use = setdiff(object@cell.names,cells.1))
    } else {
      cells.2 <- Seurat::WhichCells(object = object, ident = ident.2)
    }
  }
  cells.2 <- setdiff(x = cells.2, y = cells.1)
  
    if (max.cells.per.ident < Inf) {
    set.seed(seed = random.seed)
    if (length(cells.1) > max.cells.per.ident) cells.1 = sample(x = cells.1, size = max.cells.per.ident)
    if (length(cells.2) > max.cells.per.ident) cells.2 = sample(x = cells.2, size = max.cells.per.ident)
    }
  
  # gene selection (based on percent expressed)
  thresh.min <- 0
  data.temp1 <- round(
    x = apply(
      X = data.use[genes.use, cells.1, drop = F],
      MARGIN = 1,
      FUN = function(x) {
        return(sum(x > thresh.min) / length(x = x))
        # return(length(x = x[x>thresh.min]) / length(x = x))
      }
    ),
    digits = 3
  )
  
  data.temp2 <- round(
    x = apply(
      X = data.use[genes.use, cells.2, drop = F],
      MARGIN = 1,
      FUN = function(x) {
        return(sum(x > thresh.min) / length(x = x))
        # return(length(x = x[x > thresh.min]) / length(x = x))
      }
    ),
    digits = 3
  )
  
  data.alpha <- cbind(data.temp1, data.temp2)
  colnames(x = data.alpha) <- c("pct.1","pct.2")
  alpha.min <- apply(X = data.alpha, MARGIN = 1, FUN = max)
  names(x = alpha.min) <- rownames(x = data.alpha)
  genes.use <- names(x = which(x = alpha.min > min.pct))
  if (length(x = genes.use) == 0) {
    stop("No genes pass min.pct threshold")
  }
  alpha.diff <- alpha.min - apply(X = data.alpha, MARGIN = 1, FUN = min)
  genes.use <- names(
    x = which(x = alpha.min > min.pct & alpha.diff > min.diff.pct)
  )
  if (length(x = genes.use) == 0) {
    stop("No genes pass min.diff.pct threshold")
  }


  #gene selection (based on average difference)
  data.1 <- apply(X = data.use[genes.use, cells.1, drop = F], MARGIN = 1, FUN = function(x) log(x = mean(x = expm1(x = x)) + pseudocount.use))
  data.2 <- apply(X = data.use[genes.use, cells.2, drop = F], MARGIN = 1, FUN = function(x) log(x = mean(x = expm1(x = x)) + pseudocount.use))
  total.diff <- (data.1 - data.2)
  if (!only.pos) genes.diff <- names(x = which(x = abs(x = total.diff) > logfc.threshold))
  if (only.pos) genes.diff <- names(x = which(x = total.diff > logfc.threshold))
  genes.use <- intersect(x = genes.use, y = genes.diff)
  if (length(x = genes.use) == 0) {
    stop("No genes pass logfc.threshold threshold")
  }
  if (test.use == "wilcox") {
    to.return <- Seurat::WilcoxDETest(
      object = object,
      assay.type = assay.type,
      cells.1 = cells.1,
      cells.2 = cells.2,
      genes.use = genes.use,
      print.bar = print.bar,
      ...
    )
  }
  
  #return results
  to.return[, "avg_logFC"] <- total.diff[rownames(x = to.return)]
  to.return <- cbind(to.return, data.alpha[rownames(x = to.return), ])
  to.return$p_val_adj = p.adjust(
    p = to.return$p_val,method = "bonferroni",
    n = nrow(x = Seurat::GetAssayData(
      object = object,
      assay.type = assay.type,
      slot = "data"
    ))
  )
  if (test.use == "roc") {
    to.return <- to.return[order(-to.return$power, -to.return$avg_logFC), ]
  } else {
    to.return <- to.return[order(to.return$p_val, -to.return$avg_logFC), ]
  }
  if (only.pos) {
    to.return <- subset(x = to.return, subset = avg_logFC > 0)
  }
  return(to.return)
}

```

```{r}
mk_markers <- CustomFindMarkers(sobj, ident.1= "15", max.cells.per.ident = 10000)
mk_markers <- mk_markers %>% 
  tibble::rownames_to_column("gene")
write_tsv(mk_markers, "mk_markers_pbmc68k.txt")
```



```{r, eval = F}


average_expr_per_cluster <- AverageExpression(sobj)
average_expr_per_cluster <- average_expr_per_cluster %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column("gene")

write_tsv(average_expr_per_cluster, "avg_expr_pbmc68k.txt")
```

```{r downsampling, message = F}
mks <- rownames(sobj@meta.data[sobj@meta.data$res.1.2 == "15", ])
not_mks <- rownames(sobj@meta.data[sobj@meta.data$res.1.2 != "15", ])

n_mks_subsampled <- seq(1,length(mks), 5)
mks_subsampled <- map(n_mks_subsampled, ~sample(mks, .x, replace = F))
all_subsampled <- map(mks_subsampled, ~c(.x, not_mks))

all_markers <- list()
for (i in seq_along(all_subsampled)){
  message(paste0("starting iteration #", i))
  tmp_dat <- SubsetData(sobj, all_subsampled[[i]])
  all_markers[[i]] <- CustomFindMarkers(tmp_dat, 
                                      ident.1 = "15", 
                                      max.cells.per.ident = 10000)
  message(paste0("done with iteration #", i))
  rm(tmp_dat)
  gc()
}

saveRDS(all_markers, "all_markers.rds")
```


```{r read_markers}
all_markers <- readRDS("all_markers.rds")
```
